Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-04-28 doc/thesis/conclusion.tex
Added gamma things

Although phylodynamic methods have been developed to fit a wide variety of
epidemiological models to phylogenetic data~\autocite{volz2012complex,
rasmussen2014phylodynamic}, our method is the first (to our knowledge) which
can fit models that do not assume panmixia. We believe this capability will be
of broad interest to the molecular evolution and epidemiology community, as it
widens the field of epidemiological parameters which may be investigated
through viral sequence data. In addition, the characterization of local contact
networks could be valuable from a public health perspective, such as for
investigating optimal vaccination strategies. This information could assist in
curtailing current epimemics, as well as preventing future epidemics of
different diseases over the same contact network.

>>>>

Although phylodynamic methods have been developed to fit a wide variety of
epidemiological models to phylogenetic data~\autocite{volz2012complex,
rasmussen2014phylodynamic}, our method is the first (to our knowledge) which
can fit models that do not assume panmixia. We believe this capability will be
of broad interest to the molecular evolution and epidemiology community, as it
widens the field of epidemiological parameters which may be investigated
through viral sequence data. In addition, the characterization of local contact
networks could be valuable from a public health perspective, such as for
investigating optimal vaccination strategies~\autocite{keeling2005networks,
peng2013vaccination, ma2013importance, rushmore2014network}. This information
could assist in curtailing current epimemics, as well as preventing future
epidemics of different diseases over the same contact network.

The paricular model we have investigated uses a preferential attachment
mechanism to generate scale-free networks resembling real-world social and
sexual networks~\autocite{liljeros2001web, schneeberger2004scale,
colgate1989risk}. Of the four parameters we considered, the pereferential
attachment power $\alpha$ was the most readily estimable. Estimating $\alpha$
with traditional epidemiological methods is challenging due to the requirement
of sampling the high-degree nodes making up the tail of the power law
distribution, although approaches such as respondent-driven
sampling~\autocite{heckathorn1997respondent} may be effective.

In closing, \software{netabc} combines phylodynamics, contact network
epidemiology, approximate Bayesian computation, and sequentiial Monte Carlo 
to provide a source of insight into network structures complementary to
traditional epidemiology.

### 2016-04-28 doc/thesis/results.Rnw
Added gamma things

>>>>

To make our analyses comparable to the existing network literature, we
estimated values of the power law exponent $\gamma$ for each of the datasets 
investigated. The results are summarized in \cref{tab:gamma}. All of the
estimated exponents were in the range $2 \leq \gamma \leq 4$ commonly reported
in the literature~\autocite{liljeros2001web, schneeberger2004scale,
colgate1989risk, brown2011transmission}. Higher estimated values of $\alpha$
translated into higher estimated values of $\gamma$, which is consistent with
the observed relationship between these two quantities (\cref{fig:gamma}).

\begin{table}
    \caption[
        Estimated power law exponents for six HIV datasets based on maximum
        \textit{a priori} estimates of BA model parameters.
    ]{
        Estimated power law exponents for six HIV datasets based on maximum
        \textit{a priori} estimates of BA model parameters. 100 networks were
        simulated using \textit{MAP} parameter estimates obtained with
        \software{netabc}. The power law exponent $\gamma$ was estimated for
        each, and the median of those estimates was used as a point estimate
        for the corresponding dataset.
    }
    \label{tab:gamma}
\end{table}
